### 2021-03-17 papers/intro_paper/discussion.tex
changes in Sec. V

\ab{We built a parameterized IMR waveform model, $\pSEOB$, that can 
measure the QNM complex frequencies of the remnant object formed \ab{through 
the merger of BHs with aligned or antialigned spins, thus extending  
previous work, which was limited} to non-spinning BHs~\cite{Brito:2018rfr}. 
The $\pSEOB$ model was recently used to infer the QNMs of some of the 
BBH's remnants detected by LIGO and Virgo during O3a~\cite{Abbott:2020jks}.}

We extended these results by analyzing GW events from the first and
second LIGO-Virgo observing runs and updated the joint measurements on
the fractional deviations in the frequency and damping time of the
least-damped QNM. As expected, the event providing the best constraint
on the least-damped QNM to be date continues to be GW150914 [see
\eqref{GW150914_delta}]. In addition, even in the most agnostic and
conservative scenario where we combine the information from different
events using an hierarchical approach, at $90\%$ credibility, we
constrain the frequency of the least damped QNM to be within $\sim
9\%$ of the GR prediction -- an improvement of a factor of $\sim 4$
over the results reported in~\cite{Abbott:2020jks}. The constraints we
obtain on the frequencies and damping time of the least damped QNM are
compatible with the ones recently reported in
Ref.~\cite{Carullo:2021dui}, where measurements on QNMs were done
using only the post-merger part of the signal was analyzed.

These constraints could in principle be used to constrain specific
non-GR theories and exotic compact objects (see
e.g.~\cite{Glampedakis:2017cgd,Cardoso:2019rvt,Maggio:2020jml}). However,
QNM computations in beyond-GR theories have only been done in an
handful of cases, mostly focusing on non-rotating or slowly-rotating
BH solutions (see
e.g.~\cite{Ferrari:2000ep,Molina:2010fb,Pani:2009wy,Blazquez-Salcedo:2016enn,Blazquez-Salcedo:2017txk,Brito:2018hjh,Franciolini:2018uyq,Cardoso:2018ptl,Tattersall:2018nve,Tattersall:2019nmh,Blazquez-Salcedo:2019nwd,Silva:2019scu,Glampedakis:2019dqh,Blazquez-Salcedo:2020jee,Blazquez-Salcedo:2020caw,Cano:2020cao})
or relying on the eikonal/geometric optics approximation to obtain
estimates of the QNMs for spinning
BHs~\cite{Blazquez-Salcedo:2016enn,Glampedakis:2017dvb,Jai-akson:2017ldo}. The
only exceptions to this rule, that we are aware of, are computations
of the QNM of Kerr-Newman BHs in Einstein-Maxwell
theory~\cite{Pani:2013ija,Pani:2013wsa,Mark:2014aja,Dias:2015wqa} and
estimates of the BBH ringdown in beyond-GR theories obtained through a
limited number of numerical
simulations~\cite{Okounkova:2019dfo,Okounkova:2019zjf}. Given these
limitations, our ability of going beyond a null test of GR and use our
results to impose precise constraints on beyond-GR theories with QNM
measurements is at this moment quite limited.

Despite these limitations there has been some recent effort to develop
parametrizations that could help mapping measurements of the
parameters $(\delta f_{\ell m 0}, \delta \tau_{\ell m0})$ onto
constraints to specific beyond-GR theories. This includes for example
the parametrization proposed in Ref.~\cite{Maselli:2019mjd}, recently
applied to LIGO-Virgo GW events in Ref. ~\cite{Carullo:2021dui}, where
deviations from the GR QNMs explicitly depend on a perturbative
expansion in the BH spin and possible extra non-GR parameters, or the
proposal of Refs.~\cite{Cardoso:2019mqo,McManus:2019ulj} where
deviations from the GR QNMs are mapped onto generic small
modifications of the perturbation equations describing the QNMs. Other
examples also include proposals to map deviations from the GR QNMs to
coefficients in generic effective field theory
actions~\cite{Cardoso:2018ptl,Franciolini:2018uyq,Cano:2020cao} or to
directly relate measurements of the QNM complex frequencies to a
parametrized non-GR BH
metric~\cite{Glampedakis:2017dvb,Suvorov:2021amy,Volkel:2020daa} which
could be then used jointly with measurements coming from the Event
Horizon Telescope to obtain stronger constraints on deviations from
GR~\cite{Volkel:2020daa,Volkel:2020xlc,Psaltis:2020lvx,Yang:2021zqy}.
We should note, however, that all these parametrizations are either
limited to non-spinning BHs or make use of a series expansion in the
BH spin which might limit their accuracy for highly-spinning BHs.

In the future, it would also be important to test whether this model
could be used to detect deviations in waveforms obtained through
numerical relativity simulations of specific beyond-GR theories. Such
simulations are still at their infancy and have so far only been done
for a handful of theories, focusing mostly on proof-of-concept
simulations~\cite{Healy:2011ef,Berti:2013gfa,Cao:2013osa,Okounkova:2017yby,Hirschmann:2017psw,Witek:2018dmd,Okounkova:2019dfo,Okounkova:2019zjf,Okounkova:2020rqw,East:2020hgw}. Nonetheless,
given the recent efforts put forward in order to simulate BBHs in
beyond-GR theories, we hope that accurate non-GR IMR waveforms will
become available in the near future.

Finally, an obvious generalization of this work would be to extend the
parameterized IMR model to generic precessing BBHs, which can in
principle be easily done using the recently developed multipolar EOB
model reported in Ref.~\cite{Ossokine:2020kjp}.